Machine Learning Tools for Open Cluster Characterization with Gaia DR2 Data

### 2021-02-06 paper/paper.tex
paper: Update Introduction
Update Introduction with suggestions made by @cguz

Co-Authored-By: Cesar Guzmann <15159632+cguz@users.noreply.github.com>

\section{INTRODUCTION}

Stellar open clusters (OCs) \cite{janes1982open} are groups of stars
gravitationally bounded originated from a single molecular gas cloud.
They share the same chemical composition and age. Moreover,
they have similar relative positions and proper motion.
Those astronomical objects are relevant to understand the spiral structure,
the dynamics and the chemical evolution of our galaxy.

The study of OCs has been pushed forward thanks to the huge and precise dataset
from the Gaia mission Gaia DR2 available since 2018
\cite{collaboration2016description} and \cite{gaia2018gaia}.
This dataset has helped to review already known open clusters and to find new ones.

Stars that belong to the same OC share relative positions,
inherited from their original gas cloud.
This means that their distances to the Earth are similar for all of them and,
therefore, they have a narrow dispersion in their parallax value.
They also share similar values of proper motion, both in right ascension and declination.
Another property they share is their chemical composition.
Thus, their metallicity must be uniform,
since these stars were born from the same gas cloud and at the same time stage.
However, to take this last property into account, we are faced with the drawback
that this parameter is poorly reported in Gaia DR2 database.

Normally, the characterization process takes into account astrometric features
such as proper motion in right ascension and declination, and parallax.
Photometric features like $E(b-r) / G_{mag}$ are considered too.
These last variables are used for generating the H-R diagram of the OC candidate stars.
This stars should present a sharp profile corresponding to an isochrone curve derived
from a theoretical model.
This model does consider metallicity, mass and brightness of the stars involved.
At the moment, a set of tools have been used for this purpose, as explained later.
But in general, these tools require supervised and parameterized models.
Therefore, a previous knowledge of the cluster is necessary or the process must be
repeated iteratively in order to achieve valid results.

Our aim is to build a machine learning model capable of characterizing an open
cluster inside a stellar region with knowing nothing about the region.
This model will take advantage of these features to train an Artificial Neural
Network (ANN) that will clusterize all stars within the given region.
One of these groups will be the open cluster that interests us.
With that purpose we have adapted the Unsupervised Deep Embedding for Clustering
Analysis \cite{xie2016unsupervised} to our problem.

This model fulfills one of our main requirements.
It has to be \emph{non-supervised} and \emph{non-parameterized} in order to make easier
the automation process of analyzing a wide range of regions with different typologies.
Another important requirement for our model is that it must be computationally efficient
in order to run in common workstations.
The model has been developed with Python using the Keras framework which takes
advantage of modern GPUs to make its computations.
That increases significantly the computational capacity of our model making possible
to run it on regular workstations with one of those GPUs.

Since the aim of this work is not to find new OCs but to improve the characterization
of the already known ones, we have taken the approach of analyzing known regions.
These regions have been taken from the OpenClust catalogue \cite{dias2002new}
(see Figure \ref{fig:openclust_catalogue}) and downloaded from Gaia DR2 database
increasing the radius of each region by a factor $1.5$.

\begin{figure}
    \includegraphics[width=\columnwidth]{../figures/openclust_catalogue.png}
    \caption{OpenClust Catalogue Distribution}
    \label{fig:openclust_catalogue}
\end{figure}

Finally, we have compared the results obtained with the model we purposed with another
method that involves tools from the Virtual Observatory (VO) such as Clusterix \cite{balaguer2020clusterix}.

As shown in Section \ref{sec:results}, the results obtained are promising.
This allows us to consider our model as a valid tool that could be part
of the standard set of tools for open clusters characterization.

\section{STATE OF THE ART}

@@

\section{AIMS AND METHOD}

>>>>

\section{Introduction}

Stellar OCs~\cite{janes1982open} are groups of stars gravitationally
bounded originated from a single molecular gas cloud.
They share the same chemical composition and age, and that is the reason
that their metallicity must be uniform since those stars were born from
the same gas cloud and at the same time stage. Likewise, they have similar
relative positions inherited from their original gas cloud, which means that
their distances to the Earth are the same. Therefore, they have a narrow
dispersion in their parallax value. Moreover, they share similar values
of proper motion, both in right ascension and declination. For all these
properties, we can see why the stellar OCs are relevant to understand the
spiral structure, dynamics, and chemical evolution of our galaxy.

The study of OCs advanced thanks to the huge and precise dataset
from the Gaia DR2 mission, available since 2018 from
\cite{collaboration2016description} and \cite{gaia2018gaia}.
Although, the Gaia DR2 database misreported the chemical composition
of some OCs, the database has helped to review already known OCs and
to find new ones.

The characterization of an OC refers to the statistical determination
of the stars that form it. Therefore, the characterization is based on
the probability of a star to be a member of the cluster~\cite{sampedro2016caracterizacion}.

Normally, the characterization process considers astrometic features such as
proper motion in right ascension and declination, or parallax.
In some cases~\cite{oliveira2013fitting}, it also considers photometric features,
which helps to generate the H-R diagram~\cite{hypki2018gaia} of the OC candidate stars.
Those OC candidates should present a sharp profile corresponding to an isochrone curve
derived from a theoretical model. The model considers metallicity, mass and brightness
of the stars involved. For all this purpose, we used a set of tools that require
supervised and parameterized models. It means that we need a previous knowledge of the
cluster, otherwise we must repeat the process iteratively to achieve valid results.

In the present paper, we propose a machine learning model capable of characterizing
OCs inside a stellar region with no previous knowledge about the region.
Our model takes advantage of some features such as proper motion in right ascension
and declination, and parallax to train an Artificial Neural Network.
The ANN clusters all the stars within given region in groups.
One of those groups will be the open cluster that interests us.

Our model contributes with several novelties: i) it is \emph{non-supervised} and
\emph{non-parameterized}, making easier the automation process of analyzing a
wide range of regions with different typologies, and ii) it is computationally
efficient to run in common workstations because it has been developed with
Python using the Keras framework which takes advantage of modern GPUs to perform
its computations. That increases significantly the computational capacity of
our model doing it possible to run on regular workstations.

This paper is organized as follows. Section~\ref{sec:related_works} presents
some related works and Section~\ref{sec:aims_and_method} describes in more detail
our model. The following Section~\ref{sec:results}, shows the results and compares
them with a method that involves tools from the Virtual Observatory (VO) such as
Clusterix~\cite{balaguer2020clusterix}. Finally, the last section concludes and
outlines future research lines.

\section{Related works}
\label{sec:related_works}

@@

\section{Open Cluster Characterization method}
\label{sec:aims_and_method}
